---
title: "Report"
author: "TBD"
date: "`r Sys.Date()`"
output: html_document
---

This chunk is to refresh R studio and library the required packages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("./scripts/setup.r")
```

Now the initial loan data will be loaded into R.

```{r data, include=TRUE}
all_loan_data <- read.csv(file = here::here('datasets', 'all_loan_data.csv')) # Load in the loan data
ext_loan_data <- read.csv(file = here::here('datasets', 'extended_loan_data.csv'))  # Load in the extended loan data with time and regions
```

Since this model is made to predict the chance of a given customer defaulting at the time of application. Many variables are not applicable. These include:

-   "credit_age_yrs"
-   "delinq_2yrs"
-   "dti"
-   "inq_last_6mths"
-   "last_pymnt_amnt"
-   "pub_rec"
-   "open_acc"
-   "total_acc"
-   "total_int"
-   "total-prncp"
-   "revol_bal"
-   "revol_util"

Thus removing these covariates leaves us with 5 categorical variables and 3 continuous variables.

```{r}

# Make all categorical variables into factors for R 
#(hashed out counts just in case we do want some counts as categorical variables)

all_cat_var <- c("term",
                 "emp_length",
                 "home_ownership", 
                 "verification_status", 
                 "purpose")

all_loan_data[,all_cat_var] <- lapply(all_loan_data[all_cat_var], factor)
lapply(all_loan_data[all_cat_var],levels)
```

Let's take a look at the data.

```{r sum, include=TRUE}
summary(all_loan_data) # Summary of data
```

```{r explore, include=TRUE}
### Make plots for the 5 categorical variables vs repay failure 
## Plot the proportion and log proportion of failure for term
# Obtain number of people in each term

termsum <- count(all_loan_data, term)

# Obtain number of people that failed in each term
loan_data_fail <- filter(all_loan_data, repay_fail == "1")
termfailsum <- count(loan_data_fail, term)

# Obtain proportion and log(proportion) of failure in each term
PropM36F <- termfailsum$n[1] / termsum$n[1]
PropM60F <- termfailsum$n[1] / termsum$n[2]

# Create data frame containing proportion and log(proportion) of failure in each term
propterm <- c(PropM36F, PropM60F)
logpropterm <- c(logPropM36F, logPropM60F)
proptermdf <- as.table(propterm)
proptermdf <- as.data.frame(proptermdf)
proptermdf <- cbind(proptermdf, logpropterm)

# Graph the data

ptermfail <- ggplot(proptermdf, aes(x=Var1, y=Freq)) + 
                geom_dotplot(binaxis='y', stackdir='center') +
                scale_x_discrete(labels = term_labels) +
                labs(x = "Term",y = "Repay Failure Proportion") 

logptermfail <- ggplot(proptermdf, aes(x=Var1, y=logpropterm)) + 
                    geom_dotplot(binaxis='y', stackdir='center') +
                    scale_x_discrete(labels = term_labels) +
                    labs(x = "Term",y = "Repay Failure Log(Proportion)") 

ggarrange(ptermfail,logptermfail)
```


```{r}
## Plot the proportion and log proportion of failure for emp_length

# Obtain number of people in each emp_length
empsum <- count(all_loan_data, emp_length)

# Obtain number of people that failed in each emp_length
empfailsum <- count(loan_data_fail, emp_length)

# Obtain proportion and log(proportion) of failure in each emp_length
prop_emp_length <- empfailsum[,2] / empsum[,2]

# Obtain proportion and log(proportion) of failure in each emp_length
prop_emp_length_log <- sapply(prop_emp_length, log)
propempdf <- data.frame(emp_length = empsum$emp_length, frequency = prop_emp_length, log_frequency = prop_emp_length_log )

pempfail <- ggplot(propempdf, aes(x=emp_length, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propempdf$emp_length) +
  labs(x = "Emp_Length",y = "Repay Failure Proportion") 

logpempfail <- ggplot(propempdf, aes(x=emp_length, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propempdf$emp_length) +
  labs(x = "Emp_length",y = "Repay Failure Log(Proportion)") 

pempfail ; logpempfail
```

```{r}
## Plot the proportion and log proportion of failure for home_ownership
# Obtain number of people in each home_ownership
homesum <- summary(All_loan_data$home_ownership)
HomeMorg <- homesum[1]
HomeNone <- homesum[2]
HomeOther <- homesum[3]
HomeOwn <- homesum[4]
HomeRent <- homesum[5]
# Obtain number of people that failed in each home_ownership
homefail <- Loan_data_fail$home_ownership
homefailsum <- summary(homefail)
HomeMorgF <- homefailsum[1]
HomeNoneF <- homefailsum[2]
HomeOtherF <- homefailsum[3]
HomeOwnF <- homefailsum[4]
HomeRentF <- homefailsum[5]
# Obtain proportion and log(proportion) of failure in each home_ownership
PropHomeMorgF <- HomeMorgF / HomeMorg
PropHomeNoneF <- HomeNoneF / HomeNone
PropHomeOtherF <- HomeOtherF / HomeOther
PropHomeOwnF <- HomeOwnF / HomeOwn
PropHomeRentF <- HomeRentF / HomeRent
LogPropHomeMorgF <- log(PropHomeMorgF)
LogPropHomeNoneF <- log(PropHomeNoneF)
LogPropHomeOtherF <- log(PropHomeOtherF)
LogPropHomeOwnF <- log(PropHomeOwnF)
LogPropHomeRentF <- log(PropHomeRentF)
# Create data frame containing proportion and log(proportion) of failure in each home_ownership
prophome <- c(PropHomeMorgF, PropHomeNoneF, PropHomeOtherF, PropHomeOwnF, PropHomeRentF)
logprophome <- c(LogPropHomeMorgF, LogPropHomeNoneF, LogPropHomeOtherF, LogPropHomeOwnF, LogPropHomeRentF)
prophomedf <- as.table(prophome)
prophomedf <- as.data.frame(prophomedf)
prophomedf <- cbind(prophomedf, logprophome)
# Graph the data
home_labels <- c("Mortgage", "None", "Other", "Own", "Rent") # Make labels for the box plots with home_ownership as a variable
phomefail<-ggplot(prophomedf, aes(x=Var1, y=Freq)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = home_labels) +
  labs(x = "Home_Ownership",y = "Repay Failure Proportion") 
phomefail
logphomefail<-ggplot(prophomedf, aes(x=Var1, y=logprophome)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = home_labels) +
  labs(x = "Home_Ownership",y = "Repay Failure Log(Proportion)") 
logphomefail

## Plot the proportion and log proportion of failure for Verification_status
# Obtain number of people in each verification_status
verisum <- summary(All_loan_data$verification_status)
NotVeri <- verisum[1]
SouVeri <- verisum[2]
Veri <- verisum[3]
# Obtain number of people that failed in each Verification_status
verifail <- Loan_data_fail$verification_status
verifailsum <- summary(verifail)
NotVeriF <- verifailsum[1]
SouVeriF <- verifailsum[2]
VeriF <- verifailsum[3]
# Obtain proportion and log(proportion) of failure in each Verification_status
PropNotVeriF <- NotVeriF / NotVeri
PropSouVeriF <- SouVeriF / SouVeri
PropVeriF <- VeriF / Veri
LogPropNotVeriF <- log(PropNotVeriF)
LogPropSouVeriF <- log(PropSouVeriF)
LogPropVeriF <- log(PropVeriF)
# Create data frame containing proportion and log(proportion) of failure in each Verification_status
propveri <- c(PropNotVeriF, PropSouVeriF, PropVeriF)
logpropveri <- c(LogPropNotVeriF, LogPropSouVeriF, LogPropVeriF)
propveridf <- as.table(propveri)
propveridf <- as.data.frame(propveridf)
propveridf <- cbind(propveridf, logpropveri)
# Graph the data
veri_labels <- c("Not Verified", "Source Verified", "Verified") # Make labels for the box plots with Verification_status as a variable
pverifail<-ggplot(propveridf, aes(x=Var1, y=Freq)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = veri_labels) +
  labs(x = "Verification_Status",y = "Repay Failure Proportion") 
pverifail
logpverifail<-ggplot(propveridf, aes(x=Var1, y=logpropveri)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = veri_labels) +
  labs(x = "Verification_Status",y = "Repay Failure Log(Proportion)") 
logpverifail

## Plot the proportion and log proportion of failure for purpose
# Obtain number of people in each purpose
purpsum <- summary(All_loan_data$purpose)
Car <- purpsum[1]
Cred <- purpsum[2]
Dept <- purpsum[3]
Edu <- purpsum[4]
Home <- purpsum[5]
House <- purpsum[6]
Maj <- purpsum[7]
Med <- purpsum[8]
Move <- purpsum[9]
Oth <- purpsum[10]
Renew <- purpsum[11]
Small <- purpsum[12]
Vac <- purpsum[13]
Wed <- purpsum[14]
# Obtain number of people that failed in each purpose
purpfail <- Loan_data_fail$purpose
purpfailsum <- summary(purpfail)
CarF <- purpfailsum[1]
CredF <- purpfailsum[2]
DeptF <- purpfailsum[3]
EduF <- purpfailsum[4]
HomeF <- purpfailsum[5]
HouseF <- purpfailsum[6]
MajF <- purpfailsum[7]
MedF <- purpfailsum[8]
MoveF <- purpfailsum[9]
OthF <- purpfailsum[10]
RenewF <- purpfailsum[11]
SmallF <- purpfailsum[12]
VacF <- purpfailsum[13]
WedF <- purpfailsum[14]
# Obtain proportion and log(proportion) of failure in each purpose
PropCarF <- CarF / Car
PropCredF <- CredF / Cred
PropDeptF <- DeptF / Dept
PropEduF <- EduF / Edu
PropHomeF <- HomeF / Home
PropHouseF <- HouseF / House
PropMajF <- MajF / Maj
PropMedF <- MedF / Med
PropMoveF <- MoveF / Move
PropOthF <- OthF / Oth
PropRenewF <- RenewF / Renew
PropSmallF <- SmallF / Small
PropVacF <- VacF / Vac
PropWedF <- WedF / Wed
LogPropCarF <- log(PropCarF)
LogPropCredF <- log(PropCredF)
LogPropDeptF <- log(PropDeptF)
LogPropEduF <- log(PropEduF)
LogPropHomeF <- log(PropHomeF)
LogPropHouseF <- log(PropHouseF)
LogPropMajF <- log(PropMajF)
LogPropMedF <- log(PropMedF)
LogPropMoveF <- log(PropMoveF )
LogPropOthF <- log(PropOthF)
LogPropRenewF <- log(PropRenewF)
LogPropSmallF <- log(PropSmallF)
LogPropVacF <- log(PropVacF)
LogPropWedF <- log(PropWedF)
# Create data frame containing proportion and log(proportion) of failure in each purpose
proppurp <- c(PropCarF,PropCredF,PropDeptF,PropEduF,PropHomeF,PropHouseF,PropMajF,PropMedF,PropMoveF,PropOthF,PropRenewF,PropSmallF,PropVacF,PropWedF)
logproppurp <- c(LogPropCarF,LogPropCredF,LogPropDeptF,LogPropEduF,LogPropHomeF,LogPropHouseF,LogPropMajF,LogPropMedF,LogPropMoveF,LogPropOthF,LogPropRenewF,LogPropSmallF,LogPropVacF,LogPropWedF)
proppurpdf <- as.table(proppurp)
proppurpdf <- as.data.frame(proppurpdf)
proppurpdf <- cbind(proppurpdf, logproppurp)
# Graph the data
purp_labels <- c("Car", "Credit_card", "Dept_Consilidation","Educational","Home_Improvement","House","Major_Purchase","Medical","Moving","Other","Renewable_energy","Small_Business","Vacation","Wedding") # Make labels for the box plots with purpose as a variable
ppurpfail<-ggplot(proppurpdf, aes(x=Var1, y=Freq)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = purp_labels) +
  labs(x = "Purpose",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))
ppurpfail
logppurpfail<-ggplot(proppurpdf, aes(x=Var1, y=logproppurp)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = purp_labels) +
  labs(x = "Purpose",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))
logppurpfail
```

```{r explore, include=TRUE}
### Make plots for the 3 continuous variables vs repay failure and plots for colinearity
RepayFail_labels <- c("0","1") # Make labels for the box plots with repay failur of the x-axis

#Box plot for loan_amount vs repay_fail
ploanamount <- ggplot(data = All_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = loan_amnt,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Loan_amount")
ploanamount

#Box plot for int_rate vs repay_fail
pinterestrate <- ggplot(data = All_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = int_rate,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Int_rate")
pinterestrate 

#Box plot for annual_inc vs repay_fail
pannualincome <- ggplot(data = All_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = annual_inc,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Annual_inc")
pannualincome

# Test for colinarity 
cortest <- ggpairs(All_loan_data, columns = c(1, 3, 6))
cortest
```
