---
title: "Report"
author: "TBD"
date: "`r Sys.Date()`"
output: html_document
---

# Part 1 : Fixed Effects Model

This chunk is to refresh R studio and library the required packages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("./scripts/setup.r")

all_loan_data <- read.csv(file = here::here('datasets', 'all_loan_data.csv')) # Load in the loan data
app_cov <- c("loan_amnt","term","int_rate","emp_length","home_ownership","annual_inc","verification_status","purpose","repay_fail")
all_loan_data <- all_loan_data[,app_cov]
all_cat_var <- c("repay_fail",
                 "term",
                 "emp_length",
                 "home_ownership", 
                 "verification_status", 
                 "purpose")

all_loan_data[,all_cat_var] <- lapply(all_loan_data[all_cat_var], factor)
lapply(all_loan_data[all_cat_var],levels) # Make all categorical variables into factors for R 
```

Now the initial loan data will be loaded into R.


Since this model is made to predict the chance of a given customer defaulting at the time of application. Many variables are not applicable. These include:

-   "credit_age_yrs"
-   "delinq_2yrs"
-   "dti"
-   "inq_last_6mths"
-   "last_pymnt_amnt"
-   "pub_rec"
-   "open_acc"
-   "total_acc"
-   "total_rec_int"
-   "total_rec_prncp"
-   "revol_bal"
-   "revol_util"

Thus removing these covariates leaves us with 5 categorical variables and 3 continuous variables.

- 
## Data Exploration

Let's take a look at the data.

```{r sum, include=TRUE}
summary(all_loan_data) # Summary of data
```

Now we are going to make our categorical variables into proportions 

```{r}
source(here("scripts","prop_data.r"))
```

Plot these 

```{r}
graph_pempfail; graph_logpempfail ; graph_ptermfail ; graph_logptermfail ; graph_phomefail ; graph_logphomefail 

```



```{r explore, include=TRUE}
### Make plots for the 3 continuous variables vs repay failure and plots for colinearity

RepayFail_labels <- c("0","1") # Make labels for the box plots with repay failur of the x-axis

#Box plot for loan_amount vs repay_fail
ploanamount <- ggplot(data = all_loan_data,
                      mapping = aes(
                                   x = factor(repay_fail),
                                   y = loan_amnt,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Loan_amount")
ploanamount

#Box plot for int_rate vs repay_fail
pinterestrate <- ggplot(data = all_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = int_rate,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Int_rate")
pinterestrate 

#Box plot for annual_inc vs repay_fail
pannualincome <- ggplot(data = all_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = annual_inc,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Annual_inc")
pannualincome

# Test for colinarity 
cortest <- ggpairs(all_loan_data, columns = c(1, 3, 6))
cortest
```
```{r exploreall data set logistic (for 3 continous variables), include=TRUE}
### Do logistic exploration for all_loan_data
## Explore relationship between log(p/(1-p)) for loan_amnt
xa1 <- all_loan_data$loan_amnt
ya <- all_loan_data$repay_fail
ya <- as.integer(ya)
ya[ya == 1] <- 0
ya[ya == 2] <- 1
gala <- cut(xa1, breaks=quantile(xa1,seq(0,100,5)/100))
yamla <- tapply(ya, gala, mean)
xamla <- tapply(xa1, gala, mean)
yampla <- log(yamla/(1-yamla))
plot(xamla,yampla,xlab="loan_amnt")

## Explore relationship between log(p/(1-p)) for int_rate
xa2 <- all_loan_data$int_rate
gair <- cut(xa2, breaks=quantile(xa2,seq(0,100,5)/100))
yamir <- tapply(ya, gair, mean)
xamir <- tapply(xa2, gair, mean)
yampir <- log(yamir/(1-yamir))
plot(xamir,yampir,xlab="int_rate")

## Explore relationship between log(p/(1-p)) for annual_inc
xa3 <- all_loan_data$annual_inc
gaai <- cut(xa3, breaks=quantile(xa3,seq(0,100,5)/100))
yamai <- tapply(ya, gaai, mean)
xamai <- tapply(xa3, gaai, mean)
yampai <- log(yamai/(1-yamai))
plot(xamai,yampai,xlab="annual_inc")
```

# Part 2: Mixed Effects Model

```{r data, include=TRUE}
ext_loan_data <- read.csv(file = here::here('datasets', 'extended_loan_data.csv'))  # Load in the extended loan data with time and regions
## Extended Model 
```{r exploreext data set logistic (for 3 continous variables), include=TRUE}
### Do logistic exploration for ext_loan_data
## Explore relationship between log(p/(1-p)) for loan_amnt
x1 <- ext_loan_data$loan_amnt
y <- ext_loan_data$repay_fail
gla <- cut(x1, breaks=quantile(x1,seq(0,100,5)/100))
ymla <- tapply(y, gla, mean)
xmla <- tapply(x1, gla, mean)
ympla <- log(ymla/(1-ymla))
plot(xmla,ympla,xlab="loan_amnt")

## Explore relationship between log(p/(1-p)) for int_rate
x2 <- ext_loan_data$int_rate
gir <- cut(x2, breaks=quantile(x2,seq(0,100,5)/100))
ymir <- tapply(y, gir, mean)
xmir <- tapply(x2, gir, mean)
ympir <- log(ymir/(1-ymir))
plot(xmir,ympir,xlab="int_rate")

## Explore relationship between log(p/(1-p)) for annual_inc
x3 <- ext_loan_data$annual_inc
gai <- cut(x3, breaks=quantile(x3,seq(0,100,5)/100))
ymai <- tapply(y, gai, mean)
xmai <- tapply(x3, gai, mean)
ympai <- log(ymai/(1-ymai))
plot(xmai,ympai,xlab="annual_inc")
```

```{r exploreext data set 5 catagorical variables, include=TRUE}
## term

# Obtain number of people in each term
exttermsum <- count(ext_loan_data, term)

# Obtain number of people that failed in each term
ext_loan_data_fail <- filter(ext_loan_data, repay_fail == "1")
exttermfailsum <- count(ext_loan_data_fail, term)

# Obtain proportion and log(proportion) of failure in each term
extprop_term <- exttermfailsum[,2] / exttermsum[,2]
extprop_term_log <- log(extprop_term)

# Obtain proportion and log(proportion) of failure in each term
extproptermdf <- data.frame(term = exttermsum$term, frequency = extprop_term, log_frequency = extprop_term_log)

# Graph for ext term
graph_extptermfail <- ggplot(extproptermdf, aes(x=term, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extproptermdf$term) +
  labs(x = "term",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extlogptermfail <- ggplot(extproptermdf, aes(x=term, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extproptermdf$term) +
  labs(x = "term",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extptermfail ; graph_extlogptermfail

## Emp_length

# Obtain number of people in each emp_length
extempsum <- count(ext_loan_data, emp_length)

# Obtain number of people that failed in each emp_length
ext_loan_data_fail <- filter(ext_loan_data, repay_fail == "1")
extempfailsum <- count(ext_loan_data_fail, emp_length)

# Obtain proportion and log(proportion) of failure in each emp_length
extprop_emp_length <- extempfailsum[,2] / extempsum[,2]
extprop_emp_length_log <- log(extprop_emp_length)

# Obtain proportion and log(proportion) of failure in each emp_length
extpropempdf <- data.frame(emp_length = extempsum$emp_length, frequency = extprop_emp_length, log_frequency = extprop_emp_length_log)

# Graph for ext emp_length
graph_extpempfail <- ggplot(extpropempdf, aes(x=emp_length, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extpropempdf$emp_length) +
  labs(x = "Emp_Length",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extlogpempfail <- ggplot(extpropempdf, aes(x=emp_length, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extpropempdf$emp_length) +
  labs(x = "Emp_length",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extpempfail ; graph_extlogpempfail

## home_ownership

# Obtain number of people in each home_ownership
exthomesum <- count(ext_loan_data, home_ownership)

# Obtain number of people that failed in each home_ownership
ext_loan_data_fail <- filter(ext_loan_data, repay_fail == "1")
exthomefailsum <- count(ext_loan_data_fail, home_ownership)

# Obtain proportion and log(proportion) of failure in each home_ownership
extprop_home_ownership <- exthomefailsum[,2] / exthomesum[,2]
extprop_home_ownership_log <- log(extprop_home_ownership)

# Obtain proportion and log(proportion) of failure in each home_ownership
extprophomedf <- data.frame(home_ownership = exthomesum$home_ownership, frequency = extprop_home_ownership, log_frequency = extprop_home_ownership_log)

# Graph for ext home_ownership
graph_extphomefail <- ggplot(extprophomedf, aes(x=home_ownership, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extprophomedf$home_ownership) +
  labs(x = "home_ownership",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extlogphomefail <- ggplot(extprophomedf, aes(x=home_ownership, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extprophomedf$home_ownership) +
  labs(x = "home_ownership",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extphomefail ; graph_extlogphomefail

## verification_status

# Obtain number of people in each verification_status
extverisum <- count(ext_loan_data, verification_status)

# Obtain number of people that failed in each verification_status
ext_loan_data_fail <- filter(ext_loan_data, repay_fail == "1")
extverifailsum <- count(ext_loan_data_fail, verification_status)

# Obtain proportion and log(proportion) of failure in each verification_status
extprop_verification_status <- extverifailsum[,2] / extverisum[,2]
extprop_verification_status_log <- log(extprop_verification_status)

# Obtain proportion and log(proportion) of failure in each verification_status
extpropveridf <- data.frame(verification_status = extverisum$verification_status, frequency = extprop_verification_status, log_frequency = extprop_verification_status_log)

# Graph for ext verification_status
graph_extpverifail <- ggplot(extpropveridf, aes(x=verification_status, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extpropveridf$verification_status) +
  labs(x = "verification_status",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extlogpverifail <- ggplot(extpropveridf, aes(x=verification_status, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extpropveridf$verification_status) +
  labs(x = "verification_status",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extpverifail ; graph_extlogpverifail

## purpose

# Obtain number of people in each purpose
extpurpsum <- count(ext_loan_data, purpose)

# Obtain number of people that failed in each purpose
ext_loan_data_fail <- filter(ext_loan_data, repay_fail == "1")
extpurpfailsum <- count(ext_loan_data_fail, purpose)

# Obtain proportion and log(proportion) of failure in each purpose
extprop_purpose <- extpurpfailsum[,2] / extpurpsum[,2]
extprop_purpose_log <- log(extprop_purpose)

# Obtain proportion and log(proportion) of failure in each purpose
extproppurpdf <- data.frame(purpose = extpurpsum$purpose, frequency = extprop_purpose, log_frequency = extprop_purpose_log)

# Graph for ext purpose
graph_extppurpfail <- ggplot(extproppurpdf, aes(x=purpose, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extproppurpdf$purpose) +
  labs(x = "purpose",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extlogppurpfail <- ggplot(extproppurpdf, aes(x=purpose, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = extproppurpdf$purpose) +
  labs(x = "purpose",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0))

graph_extppurpfail ; graph_extlogppurpfail
```

```{r exploreext data set 3 continous variables, include=TRUE}
### Make plots for the 3 continuous variables vs repay failure and plots for colinearity in the extended data set (since there are extra observations)

RepayFail_labels <- c("0","1") # Make labels for the box plots with repay failure of the x-axis

#Box plot for loan_amount vs repay_fail for extended data set
pextloanamount <- ggplot(data = ext_loan_data,
                         mapping = aes(
                           x = factor(repay_fail),
                           y = loan_amnt,
                         )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Loan_amount")
pextloanamount

#Box plot for int_rate vs repay_fail
pextinterestrate <- ggplot(data = ext_loan_data,
                        mapping = aes(
                          x = factor(repay_fail),
                          y = int_rate,
                        )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Int_rate")
pextinterestrate 

#Box plot for annual_inc vs repay_fail
pextannualincome <- ggplot(data = ext_loan_data,
                        mapping = aes(
                          x = factor(repay_fail),
                          y = annual_inc,
                        )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Annual_inc")
pextannualincome

# Test for colinarity 
cortest <- ggpairs(ext_loan_data, columns = c(2, 4, 7))
cortest
>>>>>>> 1c965322079901b1c294b13f45750e64c79a569a
```

```{r exploreext data set issue_d, include=TRUE}
## Plot the proportion and log proportion of failure for issue_d

# Obtain number of people in each issue_d
issuedsum <- count(ext_loan_data, issue_d, .drop = FALSE)

# Obtain number of people that failed in each issue_d
ext_issued_fail <- count(ext_loan_data, issue_d, repay_fail == "1")
ext_missingrowIF <- c("Jun-2007", "TRUE", "0")
ext_issued_fail <- rbind(ext_issued_fail, ext_missingrowIF)
ext_issued_fail <- t(ext_issued_fail)
ext_issued_fail <- as.data.frame(ext_issued_fail)
ext_issued_fail <- relocate(ext_issued_fail, V110, .after = V55)
ext_issued_fail <- t(ext_issued_fail)
ext_issued_fail <- as.data.frame(ext_issued_fail)
ext_issued_fail$`repay_fail == "1"` <- as.logical(ext_issued_fail$`repay_fail == "1"`)
ext_issued_fail$n <- as.numeric(ext_issued_fail$n)
issuedfailsumext <- filter(ext_issued_fail, ext_issued_fail$`repay_fail == "1"` == "TRUE")

# Obtain proportion and log(proportion) of failure in each issue_d
prop_issue_d <- issuedfailsumext[,3] / issuedsum[,2]

# Obtain proportion and log(proportion) of failure in each issue_d
prop_issue_d_log <- sapply(prop_issue_d, log)
propissueddf <- data.frame(issue_d = issuedsum$issue_d, frequency = prop_issue_d, log_frequency = prop_issue_d_log )

pissuedfail <- ggplot(propissueddf, aes(x=issue_d, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propissueddf$issue_d) +
  labs(x = "issue_d",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -90,vjust = 0,hjust = 0))

logpissuedfail <- ggplot(propissueddf, aes(x=issue_d, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propissueddf$issue_d) +
  labs(x = "issue_d",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -90,vjust = 0,hjust = 0))

pissuedfail ; logpissuedfail

### Plot the proportion and log proportion of failure for each variable in issue_d

## Plot the logit of failure for each continous variable for issue_d

# Plot the logit of failure for each loan_amnt variable for issue_d
par(mfrow=c(3,3))
for(i in 1:9){
isd <- which(ext_loan_data$issue_d == issuedsum[i,1])
xamnisd <- ext_loan_data$loan_amnt[isd]
yisd <- ext_loan_data$repay_fail[isd]
gamnisd <- cut(xamnisd, breaks=quantile(xamnisd,seq(0,100,15)/100))
ymamnisd <- tapply(yisd, gamnisd, mean)
xmamnisd <- tapply(xamnisd, gamnisd, mean)
ympamnisd <- log(ymamnisd/(1-ymamnisd))
plot(xmamnisd,ympamnisd,main=paste("issue_d = ",issuedsum[i,1]),xlab="loan_amnt")
}

# Plot the logit of failure for each int_rate variable for issue_d
par(mfrow=c(3,3))
for(i in 1:9){
  isd <- which(ext_loan_data$issue_d == issuedsum[i,1])
  xinrisd <- ext_loan_data$int_rate[isd]
  yisd <- ext_loan_data$repay_fail[isd]
  ginrisd <- cut(xinrisd, breaks=quantile(xinrisd,seq(0,100,15)/100))
  yminrisd <- tapply(yisd, ginrisd, mean)
  xminrisd <- tapply(xinrisd, ginrisd, mean)
  ympinrisd <- log(yminrisd/(1-yminrisd))
  plot(xminrisd,ympinrisd,main=paste("issue_d = ",issuedsum[i,1]),xlab="int_rate")
}

# Plot the logit of failure for each annual_inc variable for issue_d
par(mfrow=c(3,3))
for(i in 1:9){
  isd <- which(ext_loan_data$issue_d == issuedsum[i,1])
  xaniisd <- ext_loan_data$annual_inc[isd]
  yisd <- ext_loan_data$repay_fail[isd]
  ganiisd <- cut(xaniisd, breaks=quantile(xaniisd,seq(0,100,15)/100))
  ymaniisd <- tapply(yisd, ganiisd, mean)
  xmaniisd <- tapply(xaniisd, ganiisd, mean)
  ympaniisd <- log(ymaniisd/(1-ymaniisd))
  plot(xmaniisd,ympaniisd,main=paste("issue_d = ",issuedsum[i,1]),xlab="annual_inc")
}
```

```{r exploreext data set zip_code, include=TRUE}
## Plot the proportion and log proportion of failure for zip_code

# Obtain number of people in each zip_code
zipcsum <- count(ext_loan_data, zip_code, .drop = FALSE)

# Obtain number of people that failed in each zip_code
ext_zipc_fail <- count(ext_loan_data, zip_code, repay_fail == "1", .drop = FALSE)
ext_zipc_fail <- filter(ext_zipc_fail, ext_zipc_fail$`repay_fail == "1"` == "TRUE")
ext_zipc_fail$`repay_fail == "1"` <- NULL
ext_zipc_fail2 <- anti_join(zipcsum, ext_zipc_fail, by = c("zip_code" = "zip_code"))
ext_zipc_fail2$n <- 0
ext_zipc_fail <- rbind(ext_zipc_fail, ext_zipc_fail2)
ext_zipc_fail$zip_code <- gsub("xx","",as.character(ext_zipc_fail$zip_code))
ext_zipc_fail <- ext_zipc_fail[order(as.integer(ext_zipc_fail$zip_code),decreasing = FALSE), ]
ext_zipc_fail$zip_code <- paste0(ext_zipc_fail$zip_code, 'xx')

# Obtain proportion and log(proportion) of failure in each zip_code
prop_zip_code <- ext_zipc_fail[,2] / zipcsum[,2]

# Obtain proportion and log(proportion) of failure in each zip_code
prop_zip_code_log <- sapply(prop_zip_code, log)
propzipcdf <- data.frame(zip_code = zipcsum$zip_code, frequency = prop_zip_code, log_frequency = prop_zip_code_log )

pzipcfail <- ggplot(propzipcdf, aes(x=zip_code, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propzipcdf$zip_code) +
  labs(x = "zip_code",y = "Repay Failure Proportion") 

logpzipcfail <- ggplot(propzipcdf, aes(x=zip_code, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propzipcdf$zip_code) +
  labs(x = "zip_code",y = "Repay Failure Log(Proportion)") 

pzipcfail ; logpzipcfail

# We did not explore the effects of zip code onto each variable since there 1 or very low number of observations in some zip codes
```

```{r exploreext data set addr_state, include=TRUE}
## Plot the proportion and log proportion of failure for addr_state

# Obtain number of people in each addr_state
addssum <- count(ext_loan_data, addr_state, .drop = FALSE)

# Obtain number of people that failed in each addr_state
ext_adds_fail <- count(ext_loan_data, addr_state, repay_fail == "1")
ext_missingrowAF <- c("ME", "TRUE", "0")
ext_adds_fail <- rbind(ext_adds_fail, ext_missingrowAF)
ext_adds_fail <- t(ext_adds_fail)
ext_adds_fail <- as.data.frame(ext_adds_fail)
ext_adds_fail <- relocate(ext_adds_fail, V100, .after = V43)
ext_adds_fail <- t(ext_adds_fail)
ext_adds_fail <- as.data.frame(ext_adds_fail)
ext_adds_fail$`repay_fail == "1"` <- as.logical(ext_adds_fail$`repay_fail == "1"`)
ext_adds_fail$n <- as.numeric(ext_adds_fail$n)
addsfailsumext <- filter(ext_adds_fail, ext_adds_fail$`repay_fail == "1"` == "TRUE")

# Obtain proportion and log(proportion) of failure in each addr_state
prop_addr_state <- addsfailsumext[,3] / addssum[,2]

# Obtain proportion and log(proportion) of failure in each addr_state
prop_addr_state_log <- sapply(prop_addr_state, log)
propaddsdf <- data.frame(addr_state = addssum$addr_state, frequency = prop_addr_state, log_frequency = prop_addr_state_log )

paddsfail <- ggplot(propaddsdf, aes(x=addr_state, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propaddsdf$addr_state) +
  labs(x = "addr_state",y = "Repay Failure Proportion") +
  theme(axis.text.x = element_text(angle = -90,vjust = 0,hjust = 0))

logpaddsfail <- ggplot(propaddsdf, aes(x=addr_state, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propaddsdf$addr_state) +
  labs(x = "addr_state",y = "Repay Failure Log(Proportion)") +
  theme(axis.text.x = element_text(angle = -90,vjust = 0,hjust = 0))

paddsfail ; logpaddsfail
<<<<<<< HEAD

### Plot the proportion and log proportion of failure for each variable in addr_state

## Plot the logit of failure for each continous variable for addr_state

# Plot the logit of failure for each loan_amnt variable for addr_state
par(mfrow=c(3,3))
for(i in 1:9){
  ads <- which(ext_loan_data$addr_state == addssum[i,1])
  xamnads <- ext_loan_data$loan_amnt[ads]
  yads <- ext_loan_data$repay_fail[ads]
  gamnads <- cut(xamnads, breaks=quantile(xamnads,seq(0,100,15)/100))
  ymamnads <- tapply(yads, gamnads, mean)
  xmamnads <- tapply(xamnads, gamnads, mean)
  ympamnads <- log(ymamnads/(1-ymamnads))
  plot(xmamnads,ympamnads,main=paste("addr_state = ",addssum[i,1]),xlab="loan_amnt")
}

# Plot the logit of failure for each int_rate variable for addr_state
par(mfrow=c(3,3))
for(i in 1:9){
  ads <- which(ext_loan_data$addr_state == addssum[i,1])
  xinrads <- ext_loan_data$int_rate[ads]
  yads <- ext_loan_data$repay_fail[ads]
  ginrads <- cut(xinrads, breaks=quantile(xinrads,seq(0,100,15)/100))
  yminrads <- tapply(yads, ginrads, mean)
  xminrads <- tapply(xinrads, ginrads, mean)
  ympinrads <- log(yminrads/(1-yminrads))
  plot(xminrads,ympinrads,main=paste("addr_state = ",addssum[i,1]),xlab="int_rate")
}

# Plot the logit of failure for each annual_inc variable for addr_state
par(mfrow=c(3,3))
for(i in 1:9){
  ads <- which(ext_loan_data$addr_state == addssum[i,1])
  xaniads <- ext_loan_data$annual_inc[ads]
  yads <- ext_loan_data$repay_fail[ads]
  ganiads <- cut(xaniads, breaks=quantile(xaniads,seq(0,100,15)/100))
  ymaniads <- tapply(yads, ganiads, mean)
  xmaniads <- tapply(xaniads, ganiads, mean)
  ympaniads <- log(ymaniads/(1-ymaniads))
  plot(xmaniads,ympaniads,main=paste("addr_state = ",addssum[i,1]),xlab="annual_inc")
}
```


## Model Building
```{r M-Build, message=FALSE, cache=TRUE}
drop<-c("credit_age_yrs",
        "delinq_2yrs",
        "dti",
        "inq_last_6mths",
        "last_pymnt_amnt",
        "pub_rec",
        "open_acc",
        "total_acc",
        "total_int",
        "total-prncp",
        "revol_bal",
        "revol_util",
        "total_rec_int",
        "total_rec_prncp")
#Full model with all possible interactions for backwards selection:
full_interaction_model <- glm(data = all_loan_data_edit,
                              formula = repay_fail ~ . ,
                              family = binomial)

#Model with no variables present for forwards selection:
null_model <- glm(data = all_loan_data_edit,
                  formula = repay_fail ~ 1,
                  family = binomial)

#Perform backward and forward selection:
backward_sel_model <- stepAIC(
  full_interaction_model,direction = "backward",trace = 0)
forward_sel_model <- stepAIC(
  null_model,
  scope = formula(full_interaction_model), 
  direction = "forward",
  trace = 0) ## trace = 0 prevents automatic output of stepAIC function.
forward_sel_model$formula
summary(forward_sel_model)
backward_sel_model$formula
summary(backward_sel_model)
m1<-glm(formula = repay_fail ~ int_rate + annual_inc + 
          term, 
        family = binomial, data = all_loan_data_edit)
summary(m1)
#forward_sel_residuals<-simulateResiduals(forward_sel_model)
#m1_residuals<-simulateResiduals(m1)

#Plot observed quantile versus expected quantile to assess distribution fit, and predicted value versus standardised residuals for unmodelled pattern in the residuals.
#plot(poisson_residuals)
# a boxplot totest for uniformity per box, and a test for homgeneity of variances between boxes. A positive test will be highlighted in red. Used just for interest, Also a test to check outliers,uniformtiy and dispersion

#plot(m1_residuals)
#plot_summs(forward_sel_model)
```

## Assessing model
```{r, Testing model message=FALSE, cache=TRUE}
predicted_values <- predict(forward_sel_model, type = 'response', newdata = training_loan_data)
plot(predicted_values)
predicted_values_vali <- predict(forward_sel_model, type = 'response', newdata = validation_loan_data)
plot(predicted_values_vali)
predicted_values_m1 <- predict(m1, type = 'response', newdata = training_loan_data)
plot(predicted_values_m1)
git config pull.rebase false
```
## Model performance
```{r, model-per message=FALSE, cache=TRUE}
install.packages("ROCR")
library(ROCR)
install.packages("gmodels")
library(gmodels)
pred <- prediction(predicted_values, training_loan_data$repay_fail)
roc.perf = performance(pred, measure = "tpr", x.measure = "fpr")
plot(roc.perf)
abline(a=0, b= 1)
pred_vali <- prediction(predicted_values_vali, validation_loan_data$repay_fail)
roc.perf = performance(pred_vali, measure = "tpr", x.measure = "fpr")
plot(roc.perf)
abline(a=0, b= 1)
auc.perf = performance(pred, measure = "auc")
auc.perf@y.values
auc.perf = performance(pred_vali, measure = "auc")
auc.perf@y.values
acc.perf = performance(pred, measure = "acc")
plot(acc.perf)
acc.perf_vali = performance(pred_vali, measure = "acc")
plot(acc.perf_vali)
ind = which.max( slot(acc.perf, "y.values")[[1]] )
acc = slot(acc.perf, "y.values")[[1]][ind]
cutoff = slot(acc.perf, "x.values")[[1]][ind]
print(c(accuracy= acc, cutoff = cutoff))

# Creating a confusion matrix
confusionMatrix(training_loan_data$repay_fail,pred_value_labels)

```

>>>>>>> 1809e38e0b3a3ed139267b819659e0b8c8ff31f6
