---
title: "Report"
author: "TBD"
date: "`r Sys.Date()`"
output: html_document
---

## Standard Model 

This chunk is to refresh R studio and library the required packages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("./scripts/setup.r")
```

Now the initial loan data will be loaded into R.

```{r data, include=TRUE}
all_loan_data <- read.csv(file = here::here('datasets', 'all_loan_data.csv')) # Load in the loan data
ext_loan_data <- read.csv(file = here::here('datasets', 'extended_loan_data.csv'))  # Load in the extended loan data with time and regions
```

Since this model is made to predict the chance of a given customer defaulting at the time of application. Many variables are not applicable. These include:

-   "credit_age_yrs"
-   "delinq_2yrs"
-   "dti"
-   "inq_last_6mths"
-   "last_pymnt_amnt"
-   "pub_rec"
-   "open_acc"
-   "total_acc"
-   "total_int"
-   "total-prncp"
-   "revol_bal"
-   "revol_util"

Thus removing these covariates leaves us with 5 categorical variables and 3 continuous variables.

```{r}

# Make all categorical variables into factors for R 
#(hashed out counts just in case we do want some counts as categorical variables)

all_cat_var <- c("term",
                 "emp_length",
                 "home_ownership", 
                 "verification_status", 
                 "purpose")

all_loan_data[,all_cat_var] <- lapply(all_loan_data[all_cat_var], factor)
lapply(all_loan_data[all_cat_var],levels)
```

Let's take a look at the data.

```{r sum, include=TRUE}
summary(all_loan_data) # Summary of data
```

Now we are going to make our categorical variables into proportions 

```{r}
source(here("scripts","prop_data.r"))
```

Plot these 

```{r}
graph_pempfail; graph_logpempfail ; graph_ptermfail ; graph_logptermfail ; graph_phomefail ; graph_logphomefail

```



```{r explore, include=TRUE}
### Make plots for the 3 continuous variables vs repay failure and plots for colinearity

RepayFail_labels <- c("0","1") # Make labels for the box plots with repay failur of the x-axis

#Box plot for loan_amount vs repay_fail
ploanamount <- ggplot(data = all_loan_data,
                      mapping = aes(
                                   x = factor(repay_fail),
                                   y = loan_amnt,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Loan_amount")
ploanamount

#Box plot for int_rate vs repay_fail
pinterestrate <- ggplot(data = all_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = int_rate,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Int_rate")
pinterestrate 

#Box plot for annual_inc vs repay_fail
pannualincome <- ggplot(data = all_loan_data,
       mapping = aes(
         x = factor(repay_fail),
         y = annual_inc,
       )) +
  geom_boxplot() +
  scale_x_discrete(labels = RepayFail_labels)+
  theme(axis.text.x = element_text(angle = -45,vjust = 0,hjust = 0)) +
  labs(x = "Repay Fail",y = "Annual_inc")
pannualincome

# Test for colinarity 
cortest <- ggpairs(all_loan_data, columns = c(1, 3, 6))
cortest
```


## Extended Model 

```{r exploreext data set, include=TRUE}
## Plot the proportion and log proportion of failure for issue_d

# Obtain number of people in each issue_d
issuedsum <- count(ext_loan_data, issue_d, .drop = FALSE)

# Obtain number of people that failed in each issue_d
ext_issued_fail <- count(ext_loan_data, issue_d, repay_fail == "1")
ext_missingrowIF <- c("Jun-2007", "TRUE", "0")
ext_issued_fail <- rbind(ext_issued_fail, ext_missingrowIF)
ext_issued_fail <- t(ext_issued_fail)
ext_issued_fail <- as.data.frame(ext_issued_fail)
ext_issued_fail <- relocate(ext_issued_fail, V110, .after = V55)
ext_issued_fail <- t(ext_issued_fail)
ext_issued_fail <- as.data.frame(ext_issued_fail)
ext_issued_fail$`repay_fail == "1"` <- as.logical(ext_issued_fail$`repay_fail == "1"`)
ext_issued_fail$n <- as.numeric(ext_issued_fail$n)
issuedfailsumext <- filter(ext_issued_fail, ext_issued_fail$`repay_fail == "1"` == "FALSE")

# Obtain proportion and log(proportion) of failure in each issue_d
prop_issue_d <- issuedfailsumext[,3] / issuedsum[,2]

# Obtain proportion and log(proportion) of failure in each issue_d
prop_issue_d_log <- sapply(prop_issue_d, log)
propissueddf <- data.frame(issue_d = issuedsum$issue_d, frequency = prop_issue_d, log_frequency = prop_issue_d_log )

pissuedfail <- ggplot(propissueddf, aes(x=issue_d, y=frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propissueddf$issue_d) +
  labs(x = "issue_d",y = "Repay Failure Proportion") 

logpissuedfail <- ggplot(propissueddf, aes(x=issue_d, y=log_frequency)) + 
  geom_dotplot(binaxis='y', stackdir='center') +
  scale_x_discrete(labels = propissueddf$issue_d) +
  labs(x = "issue_d",y = "Repay Failure Log(Proportion)") 

pissuedfail ; logpissuedfail
```